---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dasymetric

<!-- badges: start -->
<!-- badges: end -->

The goal of dasymetric is to display statistical data in meaningful spatial zones

## Installation

You can install the released version of dasymetric from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("dasymetric")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("JaFro96/dasymetric")
```
## Example

As a case study the districts of Münster were chosen and we try to predict the population counts on the basis of land cover data.

```{r districts, echo = FALSE}
load("data/districts.rda")
plot(sf::st_geometry(districts))
title("Districts of Münster")
```

```{r land_cover, echo = FALSE}
load("data/corine_18.rda")
plot(corine_18[1])
title("Land cover of Münster")
```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.
